---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data
```{r,echo=TRUE}
library(dslabs)
library(dplyr)
library(ggplot2)
setwd("C:/Users/KARAN/Documents/R projects")
setwd("./reproducability-week2")
data=read.csv("activity.csv")
data$date=as.Date(data$date)
```


## What is mean total number of steps taken per day?
## Mean and Toatal steps per day
```{r,echo=TRUE }
data%>%group_by(date)%>%
  summarise(totalsteps=sum(steps,na.rm=T),mean=mean(steps,na.rm=T),median=median(steps,na.rm = T))%>%
  print()%>%
  ggplot(aes(totalsteps))+
  geom_histogram()+labs(title="Sum of steps of each day")


```


## What is the average daily activity pattern?
```{r,echo=TRUE}
avgsteps=as.numeric(tapply(data$steps,data$interval,mean,na.rm=T))
intervals=unique(data$interval)
plot(intervals,avgsteps,type="l",main="Average steps for 5min interval")
```
The maximum avg steps for the interval is:
```{r,echo=TRUE}

intervals[which.max(avgsteps)]
```


## Imputing missing values
The total number of NA's in the dataset are:
```{r,echo=TRUE}
sum(is.na(data$steps))
```
Imputing the missing values by their average 5min interval and creating a new dataset without any NA values:
```{r,echo=TRUE}
data1=cbind(data,avgsteps)
data1$steps[is.na(data1$steps)]=data1$avgsteps[is.na(data1$steps)]
head(data1)


```
Histogram of the total steps per day from the new dataset and displaying the new mean and median:
```{r,echo=TRUE}
data1%>%group_by(date)%>%
  summarise(totalsteps=sum(steps,na.rm=T),mean=mean(steps,na.rm=T),median=median(steps,na.rm = T))%>%
  print()%>%
  ggplot(aes(totalsteps))+
  geom_histogram()+labs(title="Sum of steps of each day(new dataset)")


```
Comapring with the old dataset with NA values:

```{r,echo=TRUE}
data%>%group_by(date)%>%
  summarise(totalsteps=sum(steps,na.rm=T),mean=mean(steps,na.rm=T),median=median(steps,na.rm = T))%>%
  print()%>%
  ggplot(aes(totalsteps))+
  geom_histogram()+labs(title="Sum of steps of each day(old dataset)")

```
The impact of impuning missing values was huge in calculating the mean and median while the sum of steps per day remained relatively unchanged for each day
## Are there differences in activity patterns between weekdays and weekends?
###Comparing weekday and weekend trends 
```{r,echo=TRUE}
data2=data1
data2$day=weekdays(data2$date)
data2$day=ifelse(data2$day=="Saturday"|data2$day=="Sunday","weekend","weekday")
data2$day=factor(data2$day)
data2%>%group_by(interval,day)%>%summarise(avgday=mean(steps,na.rm=T))%>%ggplot(aes(x=interval,y=avgday),col=day)+geom_line()+facet_grid(day~.)+labs(x="5min interval",y="Average steps per day",main="weekdays Vs weekends")


```

